---
title: "Cleanse"
author: "Dean Stewart"
date: "14/03/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Tibbles vs dataframes

```{r page 19 examples}
my_tib <- tibble(x=1:3, y=c("blue","pink","yellow"))
my_tib
```

```{r}
my_df <- data.frame(x=1:3, y=c("blue","pink","yellow"))
my_df
```

```{r}
as_tibble(my_df)

```

```{r}
data(starwars)
starwars
```

```{r}
as.data.frame(starwars)
```

These all look the same on the screen, because rmarkdown prints them like they're a tibble anyway!
they look different in the output file


subset of a dataframe is a vector
```{r}
my_df[,1]
```

subset of a tibble is still a tibble
```{r}
my_tib[,1]
```

to extract a vector from a tibble, use the double square bracket
```{r}
my_tib[['x']]
```


## sequential variable creation

Tibbles can create variables in a sequential mannger but dataframes cannot
```{r sequential creation, echo=TRUE}
#this won't work
#anotherDF <- data.frame(nItems = c(12,45,107),
#                        cost = c(0.5,1.2,1.8),
#                        totalWorth = nItems*cost)


#this will
anothertib <- tibble(nItems = c(12,45,107),
                        cost = c(0.5,1.2,1.8),
                        totalWorth = nItems*cost)

```

