---
title: "Summarise Data"
author: "Dean Stewart"
date: "14/03/2021"
output: html_document
---

```{r setup, include=FALSE}
#ONLY TO BE RUN AFTER LOAD THE DATA SCRIPT
knitr::opts_chunk$set(echo = TRUE)
```

## Data Summaries to assess data quality

```{r summarise, echo=FALSE}
library(gapminder)
data(gapminder)
str(gapminder)
```

```{r}
class(gapminder)
```

```{r}
dim(gapminder)
```

```{r}
table(gapminder$country)
```

```{r}
prop.table(table(gapminder$year))
#gives the proportions of each result - so like table, but just divided by total count
```

```{r}
summary(gapminder)
```

```{r}
library(psych)
describe(gapminder)
```

```{r}
cor(gapminder[,3:6])
```
## Example code
from page 16 of week 3 guide

```{r examplep16, echo=TRUE}
#leave out some of the descriptive stuff done above, so start in 2nd para
library(janitor)
library(dplyr)
gapminder_adjnames <- gapminder
names(gapminder_adjnames) <- c("country", "continent", "year", "Life Exp", "population", "Gdp Per cap")
make.names(names(gapminder_adjnames))
gapminder_clean <- clean_names(gapminder_adjnames)
names(gapminder_clean)
gapminder_clean2 <- rename(gapminder_adjnames, life_exp = "Life Exp", gdp_per_cap = "Gdp Per cap")
names(gapminder_clean2)
```

```{r second part of example, echo=TRUE}
cor(gapminder[, 3:6])
cor(gapminder[, 3:6], use="pairwise.complete.obs") #use this when there are NAs in the data
```

### Other useful functions
* **nrow(dataframe)** - returns the number of rows
* **ncol(dataframe)** - returns the number of columns
* **colnames(dataframe)** or **names(dataframe)** - returns the columns *names*
* **view(dataframe)** - brings up the dataframe in a graphical table views
* **unique(vector)** - returns the unique values in a vector
* **glimpse(dataframe)** - similar to str, but easier to read


